---
title: "Citizen's Initiative Draft Code"
author: "Caitlin Dawson"
date: "15 6 2022"
output: html_document
---

# Notes

This is a draft version of code to format and process data from the Citizen's Initiative Task, meant to be combined with the full code Notebook after testing.

Notes: 
Last edited 15.6.2022 by Caitlin. Takes data downloaded from Gorilla in **blinded, semicolon separater, csv format, short form**, versions 22-25.

Version history: 
*version 23: HJ updated the experiment (a source was deleted from CI task so this version has only 5 sources. The Viiskunta source was taken down so we could not get a stable URL for it.)
*version 24: HJ updated CI task to include all 6 sources again, Viiskunta source as an image
*version 25: changes made elsewhere, CI is same as version 24

- Missing data needs to be replaced with NA
- version 22 has the full set of 6 links
- version 23 has fewer rows per ppt because of only 5 sources. All sources are links. 
- versions 24 and 25 have all sources but one source as a PDF. Find the source implementation in the column 'display'

# Extract Overall RTs

```{r Extract Overall RTs}

# Extract reaction time for the entire task to a different df
data_citiT_overallRTs <- data_citiT %>% 
  filter(Trial.Number == "END TASK") %>% 
  select(Participant.Private.ID, Reaction.Time)
names(data_citiT_overallRTs)[2] <- "overall_rt"

# Remove BEGIN TASK and END TASK rows from df; Remove the instructions/finish rows
data_citiT <- data_citiT %>% 
  filter(!Trial.Number %in% c("END TASK", "BEGIN TASK")) %>%
  filter(display != "Instructions" & display!= "Finish")

```

# Add Order Column

```{r Add Order Column}

# Baseline SAM and 4 baseline topic questions (about the petition) coded as order = 1; Rest of the trials coded based on Source_topic so that order numbers always refer to the same source
data_citiT <- data_citiT %>% 
  mutate(src_id = case_when((Source_topic == "") ~ 1,
                            (Source_topic == "Auttaa akneen, reumaan, selluliittiin ja hiustenlähtöön – mikä se on?") ~ 2,
                            (Source_topic == "Palaako peltoviljely pääosin kangasmaille?") ~ 3,
                            (Source_topic == "Punnittua puhetta turpeesta") ~ 4,
                            (Source_topic == "Suo syntyy uudestaan") ~ 5,
                            (Source_topic == "Turve on uusiutuva luonnonvara") ~ 6,
                            (Source_topic == "Turvetuotanto on ajettava alas – oikeudenmukaisesti") ~ 7))

# Create order column
order_df <- data_citiT[, c("Participant.Private.ID", "src_id")]
order_list <- list()
for (j in 1:length(unique(order_df$Participant.Private.ID))) {
  data <- order_df[which(order_df$Participant.Private.ID == unique(order_df$Participant.Private.ID)[j]), "src_id"] 
  ord <- c() 
  for (i in 1:length(data)) {
    if (all(data[i] != ord)) {
      ord <- c(ord, data[i])
    }
  }
  order_list[[j]] <- ord
}
order_vector <- c()
id <- unique(data_citiT$Participant.Private.ID)
for (i in 1:length(id)) {
  order_vector[i] <- paste(order_list[[i]], collapse = " ")
}
order_citiT <- data.frame(Participant.Private.ID = id,
                          order = order_vector)

```

# Extract Source Reading Times

```{r Extract Reading Times}

# Create subset
data_citiT_srcRTs <- data_citiT %>% 
  filter(Zone.Name == "advancementZone") %>% 
  select(Participant.Private.ID, Screen.Name, display, src_id, Reaction.Time, Zone.Name)
names(data_citiT_srcRTs)[5] <- "src_rt"

# The PDF setting gives two reaction times; Screen 1 is for the intro page and Screen 12 is the actual PDF, so we only keep Screen 12
data_citiT_srcRTs <- data_citiT_srcRTs[-(which(data_citiT_srcRTs$display == "PDF" & data_citiT_srcRTs$Screen.Name == "Screen 1" & data_citiT_srcRTs$Zone.Name == "advancementZone")), ]

# Remove two extra rows for sources 5 and 6 for one participant; replace the RTs with NAs (cannot be sure which one is the "real" one)
data_citiT_srcRTs <- data_citiT_srcRTs[-c((which(data_citiT_srcRTs$Participant.Private.ID == "5380955" & data_citiT_srcRTs$src_id == 5 & data_citiT_srcRTs$Zone.Name == "advancementZone")[1]), which(data_citiT_srcRTs$Participant.Private.ID == "5380955" & data_citiT_srcRTs$src_id == 6 & data_citiT_srcRTs$Zone.Name == "advancementZone")[1]), ]
data_citiT_srcRTs[which(data_citiT_srcRTs$Participant.Private.ID == "5380955" & data_citiT_srcRTs$src_id == 5 & data_citiT_srcRTs$Zone.Name == "advancementZone"), "src_rt"] <- NA
data_citiT_srcRTs[which(data_citiT_srcRTs$Participant.Private.ID == "5380955" & data_citiT_srcRTs$src_id == 6 & data_citiT_srcRTs$Zone.Name == "advancementZone"), "src_rt"] <- NA
data_citiT_srcRTs <- data_citiT_srcRTs[-( which(data_citiT_srcRTs$Participant.Private.ID == "5446920" & data_citiT_srcRTs$src_id == 6 & data_citiT_srcRTs$Zone.Name == "advancementZone"))[1], ]
data_citiT_srcRTs[which(data_citiT_srcRTs$Participant.Private.ID == "5446920" & data_citiT_srcRTs$src_id == 6 & data_citiT_srcRTs$Zone.Name == "advancementZone"), "src_rt"] <- NA

# Combine overall RTs and source RTs
data_citiT_RTs <- merge(data_citiT_overallRTs, data_citiT_srcRTs,
                        by = "Participant.Private.ID",
                        all = TRUE)

# Keep relevant columns
data_citiT_RTs <- data_citiT_RTs %>% 
  select(Participant.Private.ID, src_id, src_rt, overall_rt)

# One participant is missing RTs for source 6; add NAs
data_citiT_RTs[nrow(data_citiT_RTs) + 1, ] <- c(5460154, 6, NA, 1378705)

# Overall RTs
describe(data_citiT_overallRTs$overall_rt/60000) # in minutes
hist(data_citiT_overallRTs$overall_rt/60000) # in minutes
plot(sort(data_citiT_overallRTs$overall_rt/60000)) # in minutes

# Reading RTs (for all sources)
describe(data_citiT_RTs$src_rt/60000) # in minutes
hist(data_citiT_RTs$src_rt/60000,
     main = "Online source RTs",
     col = sample(colors(), 1),
     xlab = "", 
     breaks = 100)
plot(sort(data_citiT_RTs$src_rt/60000)) # in minutes

```

# Extract Likert Response Times

```{r Extract Likert Response Times}

# Extract reaction times for the Likert questions
data_citiT_likertRTs <- data_citiT %>% 
  filter(Zone.Type != "continue_button") %>% 
  select(Participant.Private.ID, src_id, Reaction.Time, Zone.Type)
# The other continue_button rows (ones where Zone.Name != advancementZone) are the same reaction times as the response_text_area rows before them; they are not needed here

# Describe
describe(data_citiT_likertRTs$Reaction.Time/1000) # in seconds
hist(data_citiT_likertRTs$Reaction.Time/1000, # in seconds
      main = "Likert RTs",
     col = sample(colors(), 1),
     xlab = "", 
     breaks = 100)
plot(sort(data_citiT_likertRTs$Reaction.Time/1000)) # in seconds

```

From Caitlin: RT data has issues for outlier removal due to the distribution, not sure yet the best method for this kind of task. We don't set an automatic upper limit and we cannot assume that the reading times represent a single cognitive process. However, I don't think we can keep all times in, since they range from one second to over an hour, with the mean around 3 minutes. The goal should be to capture as much valid data as possible, with the acknowledgement that times e.g. 1 second are valid but mean something different than longer times, e.g. someone has skimmed or skipped that source. 
* should we look into the thresholds for skimming and reading comprehension to make a point about which times we can expect to mean actually reading and which are skipping? 
* it is clear where obviously meaningless long outliers are in this data, but the question is how to justify their removal 
* can reading times be analyzed like reaction times, or is there another way? 

From Milla: I think we need separate exclusion criteria (possibly) for the overall RT (i.e., if someone went through the entire task super quickly), the reading times for each source (i.e., if someone went through a particular source very quickly), and for the Likert RTs (i.e., if someone was so quick at responsing to the questions that it can be assumed they didn't even read the question, although this might actually be true as participants might memorize the questions quickly)
* also need to think about whether we want to set maximum time limits (and if so, what these would reflect) -> NO
* can we have some missing data on a participant-level for this task (from an analysis point of view)?


# Code Source Types

```{r Code Source Types}

# Add a column to code the source authority: 1=Authority, 2=Personal
data_citiT <- data_citiT %>%
  mutate(source_authority = case_when((Source_topic == "Turvetuotanto on ajettava alas – oikeudenmukaisesti") ~ 1,
                                      (Source_topic == "Suo syntyy uudestaan") ~ 1,
                                      (Source_topic == "Palaako peltoviljely pääosin kangasmaille?") ~ 1,
                                      (Source_topic == "Punnittua puhetta turpeesta") ~ 2,
                                      (Source_topic == "Turve on uusiutuva luonnonvara") ~ 2,
                                      (Source_topic == "Auttaa akneen, reumaan, selluliittiin ja hiustenlähtöön – mikä se on?") ~ 2))

# Add a column to code the source quality: 1=Reliable, 2=Unreliable, 3=irrelevant
data_citiT <- data_citiT %>%
  mutate(source_quality = case_when((Source_topic == "Turvetuotanto on ajettava alas – oikeudenmukaisesti") ~ 1,
                                    (Source_topic == "Suo syntyy uudestaan") ~ 2,
                                    (Source_topic == "Palaako peltoviljely pääosin kangasmaille?") ~ 3,
                                    (Source_topic == "Punnittua puhetta turpeesta") ~ 1,
                                    (Source_topic == "Turve on uusiutuva luonnonvara") ~ 2,
                                    (Source_topic == "Auttaa akneen, reumaan, selluliittiin ja hiustenlähtöön – mikä se on?") ~ 3))

```


# Recode the SAM Responses as Likert Intensity and Valence Scores

Recode the SAM responses as Likert intensity and valence scores and add a line to extract the relevant rows, e.g. emotional state associated with the source topic, authority and quality. Add that to the reading time df.

```{r Recode the SAM + Petition + Source Responses}

# SAM intensity
data_citiT <- data_citiT %>%
  mutate(SAMintensity = case_when((Response == "SAM_1.1.PNG") ~ 1,
                                  (Response == "SAM_1.2.PNG") ~ 2,
                                  (Response == "SAM_1.3.PNG") ~ 3,
                                  (Response == "SAM_1.4.PNG") ~ 4,
                                  (Response == "SAM_1.5.PNG") ~ 5,
                                  (Response == "SAM_1.6.PNG") ~ 6,
                                  (Response == "SAM_1.7.PNG") ~ 7,
                                  (Response == "SAM_1.8.PNG") ~ 8,
                                  (Response == "SAM_1.9.PNG") ~ 9))

# SAM valence
data_citiT <- data_citiT %>%
  mutate(SAMvalence = case_when((Response == "SAM_2.1.PNG") ~ 1,
                                (Response == "SAM_2.2.PNG") ~ 2,
                                (Response == "SAM_2.3.PNG") ~ 3,
                                (Response == "SAM_2.4.PNG") ~ 4,
                                (Response == "SAM_2.5.PNG") ~ 5,
                                (Response == "SAM_2.6.PNG") ~ 6,
                                (Response == "SAM_2.7.PNG") ~ 7,
                                (Response == "SAM_2.8.PNG") ~ 8,
                                (Response == "SAM_2.9.PNG") ~ 9))

# Curiosity (petition)
data_citiT <- data_citiT %>%
  mutate(Curiosity = case_when((Response == "En ollenkaan uteliaasti") ~ 1,
                               (Response == "Hiukan uteliaasti") ~ 2,
                               (Response == "Jokseenkin uteliaasti") ~ 3,
                               (Response == "Hyvin uteliaasti") ~ 4,
                               (Response == "Erittäin uteliaasti") ~ 5))

# Interest (petition)
data_citiT <- data_citiT %>%
  mutate(Interest = case_when(((Response == "Ei ollenkaan kiinnostava") | 
                              (Response == "En ollenkaan kiinnostavana")) ~ 1,
                              ((Response == "Hiukan kiinnostava") | 
                              (Response == "Hiukan kiinnostavana")) ~ 2,
                              ((Response == "Jokseenkin kiinnostava") | 
                              (Response == "Jokseenkin kiinnostavana")) ~ 3,
                              ((Response == "Hyvin kiinnostava") | 
                              (Response == "Hyvin kiinnostavana")) ~ 4,
                              ((Response == "Erittäin kiinnostava") | 
                              (Response == "Erittäin kiinnostavana")) ~ 5))

# Familiar (petition)
data_citiT <- data_citiT %>%
  mutate(Familiarity = case_when((Response == "Ei ollenkaan tuttu") ~ 1,
                                 (Response == "Hiukan tuttu") ~ 2,
                                 (Response == "Jokseenkin tuttu") ~ 3,
                                 (Response == "Hyvin tuttu") ~ 4,
                                 (Response == "Erittäin tuttu") ~ 5))

# Support (petition)
data_citiT <- data_citiT %>%
  mutate(Support = case_when(((Response == "Varmasti en" & display == "Reading" & Screen.Name == "Screen 11") | (Response == "Varmasti en" & display == "Initial_assessment" & Screen.Name == "Screen 1") | (Response == "Varmasti en" & display == "PDF" & Screen.Name == "Screen 8")) ~ 1,
                                  ((Response == "Luultavasti en" & display == "Reading" & Screen.Name == "Screen 11") | (Response == "Luultavasti en" & display == "Initial_assessment" & Screen.Name == "Screen 1") | (Response == "Luultavasti en" & display == "PDF" & Screen.Name == "Screen 8")) ~ 2,
                                  ((Response == "En osaa sanoa" & display == "Reading" & Screen.Name == "Screen 11") | (Response == "En osaa sanoa" & display == "Initial_assessment" & Screen.Name == "Screen 1") | (Response == "En osaa sanoa" & display == "PDF" & Screen.Name == "Screen 8")) ~ 3,
                                  ((Response == "Luultavasti kyllä" & display == "Reading" & Screen.Name == "Screen 11") | (Response == "Luultavasti kyllä" & display == "Initial_assessment" & Screen.Name == "Screen 1") | (Response == "Luultavasti kyllä" & display == "PDF" & Screen.Name == "Screen 8")) ~ 4,
                                  ((Response == "Varmasti kyllä" & display == "Reading" & Screen.Name == "Screen 11") | (Response == "Varmasti kyllä" & display == "Initial_assessment" & Screen.Name == "Screen 1") | (Response == "Varmasti kyllä" & display == "PDF" & Screen.Name == "Screen 8")) ~ 5))

# Shareability (source)
data_citiT <- data_citiT %>%
  mutate(Shareability = case_when(((Response == "Varmasti en" & display == "Reading" & Screen.Name == "Screen 2") | (Response == "Varmasti en" & display == "PDF" & Screen.Name == "Screen 7")) ~ 1,
                                      ((Response == "Luultavasti en" & display == "Reading" & Screen.Name == "Screen 2") | (Response == "Luultavasti en" & display == "PDF" & Screen.Name == "Screen 7")) ~ 2,
                                      ((Response == "En osaa sanoa" & display == "Reading" & Screen.Name == "Screen 2") | (Response == "En osaa sanoa" & display == "PDF" & Screen.Name == "Screen 7")) ~ 3,
                                      ((Response == "Luultavasti kyllä" & display == "Reading" & Screen.Name == "Screen 2") | (Response == "Luultavasti kyllä" & display == "PDF" & Screen.Name == "Screen 7")) ~ 4,
                                      ((Response == "Varmasti kyllä" & display == "Reading" & Screen.Name == "Screen 2") | (Response == "Varmasti kyllä" & display == "PDF" & Screen.Name == "Screen 7")) ~ 5))

# Convincingness (source)
data_citiT <- data_citiT %>%
  mutate(Convincingness = case_when((Response == "Ei ollenkaan vakuuttava") ~ 1,
                                    (Response == "Ei vakuuttava") ~ 2,
                                    (Response == "Neutraali") ~ 3,
                                    (Response == "Vakuuttava") ~ 4,
                                    (Response == "Erittäin vakuuttava") ~ 5))

# Expertise (source)
data_citiT <- data_citiT %>%
  mutate(Expertise = case_when((Response == "Ei ollenkaan") ~ 1,
                               (Response == "Vähän") ~ 2,
                               (Response == "Jonkin verran") ~ 3,
                               (Response == "Paljon") ~ 4,
                               (Response == "Hyvin paljon") ~ 5))

# Reliability (source)
data_citiT <- data_citiT %>%
  mutate(Reliability = case_when((Response == "Erittäin epäluotettava") ~ 1,
                                 (Response == "Epäluotettava") ~ 2,
                                 (Response == "Ei epäluotettava eikä luotettava") ~ 3,
                                 (Response == "Luotettava") ~ 4,
                                 (Response == "Erittäin luotettava") ~ 5))

# Extract relevant information per participant per source
data_citiT_ratings <- data_citiT %>% 
  group_by(Participant.Private.ID, src_id) %>% 
  summarise(SAMintensity_final = mean(SAMintensity, na.rm = TRUE),
            SAMvalence_final = mean(SAMvalence, na.rm = TRUE),
            Curiosity_pet = mean(Curiosity, na.rm = TRUE),
            Interest_pet = mean(Interest, na.rm = TRUE),
            Familiarity_pet = mean(Familiarity, na.rm = TRUE),
            Support_pet = mean(Support, na.rm = TRUE),
            Shareability_src = mean(Shareability, na.rm = TRUE),
            Convincingness_src = mean(Convincingness, na.rm = TRUE),
            Expertise_src = mean(Expertise, na.rm = TRUE),
            Reliability_src = mean(Reliability, na.rm = TRUE))

# Add column for word count per source (I counted the headings + "normal" text but no captions or author intros etc.)
data_citiT_ratings <- data_citiT_ratings %>% 
  mutate(src_words = case_when((src_id == 1) ~ 0,
                               (src_id == 2) ~ 408,
                               (src_id == 3) ~ 327,
                               (src_id == 4) ~ 892,
                               (src_id == 5) ~ 1146,
                               (src_id == 6) ~ 422,
                               (src_id == 7) ~ 376))
data_citiT_ratings[which(data_citiT_ratings$src_words == 0), "src_words"] <- NA
# NOTE: reading speed?

# Add order column
data_citiT_final <- merge(data_citiT_ratings, order_citiT,
                          by = "Participant.Private.ID")

# Replace NaNs with NAs
data_citiT_final[is.nan(data_citiT_final)] <- NA

# Check row sum NAs
colSums(is.na(data_citiT_final))

```


```{r Combine RTs and Ratings}

# Combine
data_citiT_final <- merge(data_citiT_final, data_citiT_RTs, 
                          by = c("Participant.Private.ID", "src_id"), 
                          all = TRUE)

# Fix overall RT so that there are no NAs
data_citiT_final$overall_rt <- rep(unique(data_citiT_final$overall_rt)[which(!is.na(unique(data_citiT_final$overall_rt)))], each = 7)

# Rearrange columns
data_citiT_final <- data_citiT_final %>% 
  select(Participant.Private.ID, order, src_id, src_words, src_rt, overall_rt, Shareability_src, Convincingness_src, Expertise_src, Reliability_src, SAMintensity_final, SAMvalence_final, Curiosity_pet, Interest_pet, Familiarity_pet, Support_pet)

```

# Exclude Participants

```{r Exclude Non-Valid Participants, include = FALSE}

# Exclude nonvalid participants
data_citiT_final <- merge(data_citiT_final, data_ConsentValidity, 
                          by = "Participant.Private.ID",
                          all.x = TRUE)

# Validity/language filter; Exclude invalid participant
data_citiT_final <- data_citiT_final %>%
  filter(Validity.quantised %in% c(1, NA, "") & Language %in% c("Sujuva / äidinkieli", "Keskitaso / keskusteleva","Muu, mikä?") & Participant.Private.ID != "4887029") %>% 
  select(Participant.Private.ID, order, src_id, src_words, src_rt, overall_rt, Shareability_src, Convincingness_src, Expertise_src, Reliability_src, SAMintensity_final, SAMvalence_final, Curiosity_pet, Interest_pet, Familiarity_pet, Support_pet)

```

Potentially problematic ppts for CI: 
ri0gfe1h	5767733	skipped reading at least one article
vu1mvwkd	5691188	didn't focus on the first CI text
8zz2teyh	5629594	is dyslexic

NOTE: I didn't remove anyone so far, I think it's better to apply a threshold for when we think someone paid no attention to the source (and omit all source-related data for that participant) than handpick the participants that were honest about not concentrating.

NOTE: wide format?
