---
title: "UUT Cohen's Kappa"
author: "Milla Pihlajamaki"
date: "20 7 2021"
output: html_document
---

# File Outline

This file contains calculations for Kappa for inter-rater reliability for the Unusual Uses Task (UUT) for the ~50 pilot participants. Ratings by MP & HJ.


```{r Loading Data, include=FALSE}

#Remember to set working directory that includes the csv file!
#Session, set working directory, choose directory etc

#these packages were downloaded, found to the right under "packages"
library(vcd)
library(psych)

#for some reason the folder wasn't registered, did 1 up and not it works
UUT <- read.csv("UUT_scoring_COMBINED_REVISED_Rcsv.csv", sep = ";") 
#UUT <- read.csv("UUT_scoring_COMBINED_fixattempt_Rcsv_norow.csv", sep = ";")

#check the working directory
print(getwd())

#summary(x) - produces a summary of a given object
summary(UUT)
summary(UUT$X) #Not needed, this includes comments, removing it from df
UUT <- UUT[1:5048,1:9] #relevance?

class(UUT$PartID) #this should be factor as it is categorical technically
UUT$PartID <- factor(UUT$PartID)
head(UUT$PartID)

class(UUT$Response) #should be factor
UUT$Item <- factor(UUT$Response)
head(UUT$Response)

#Initialize variables
#UUT$S.Fluency <- NA  
#UUT$S.Flex <- NA  
#UUT$K.Fluency <- NA  
#UUT$K.Flex <- NA  

range(UUT[1:5048,4:7]) #This is correct, there are no outliers
UUT$S.Fluency <- factor(UUT$S.Fluency) #change into factor for the analysis
UUT$S.Flex <- factor(UUT$S.Flex)
UUT$K.Fluency <- factor(UUT$K.Fluency)
UUT$K.Flex <- factor(UUT$K.Flex)

summary(UUT)

```

## Data Explanation

The data look fine, we have one categorical variable for the participant ID number and one categorical variable for the item in question. We also have a character column for participants' answers. 

The important columns are columns 4 to 7, which include scores for each answer (categorical, 1=yes/accepted and 0=no/not accepted). Fluency and flexibility were rated separately, and participants got 1 point if the raters (MP & HJ) considered the answer to exhibit fluency/flexibility (separate columns for these). 


```{r Data into Contingency Table, echo=TRUE}

#We need the data in a contingency table for the analysis

#Fluency
nrow(UUT[UUT$S.Fluency == "1" & UUT$K.Fluency == "1",]) #extracting the contingencies
nrow(UUT[UUT$S.Fluency == "0" & UUT$K.Fluency == "1",])
nrow(UUT[UUT$S.Fluency == "1" & UUT$K.Fluency == "0",])
nrow(UUT[UUT$S.Fluency == "0" & UUT$K.Fluency == "0",])

UUT_Fluency <- as.table(rbind(c(nrow(UUT[UUT$S.Fluency == "1" & UUT$K.Fluency == "1",]),nrow(UUT[UUT$S.Fluency == "0" & UUT$K.Fluency == "1",])), c(nrow(UUT[UUT$S.Fluency == "1" & UUT$K.Fluency == "0",]),nrow(UUT[UUT$S.Fluency == "0" & UUT$K.Fluency == "0",]))))

categories <- c("Yes", "No")
dimnames(UUT_Fluency) <- list(K = categories, S = categories)
UUT_Fluency


#Flexibility
nrow(UUT[UUT$S.Flex == "1" & UUT$K.Flex == "1",])
nrow(UUT[UUT$S.Flex == "0" & UUT$K.Flex == "1",])
nrow(UUT[UUT$S.Flex == "1" & UUT$K.Flex == "0",])
nrow(UUT[UUT$S.Flex == "0" & UUT$K.Flex == "0",])

UUT_Flexibility <- as.table(rbind(c(nrow(UUT[UUT$S.Flex == "1" & UUT$K.Flex == "1",]),nrow(UUT[UUT$S.Flex == "0" & UUT$K.Flex == "1",])), c(nrow(UUT[UUT$S.Flex == "1" & UUT$K.Flex == "0",]),nrow(UUT[UUT$S.Flex == "0" & UUT$K.Flex == "0",]))))

categories2 <- c("Yes", "No")
dimnames(UUT_Flexibility) <- list(K = categories2, S = categories2)
UUT_Flexibility


```


## Kappa Assumptions

https://www.datanovia.com/en/lessons/cohens-kappa-in-r-for-two-categorical-variables/

Your data should met the following assumptions for computing Cohen’s Kappa.

* You have two outcome categorical variables, which can be ordinal or nominal variables.  
* The two outcome variables should have exactly the same categories.  
* You have paired observations; each subject is categorized twice by two independent raters or methods.  
* The same two raters are used for all participants.  

These apply in the present data set.



## Kappa Calculations

```{r Kappa for Fluency and Flexibility, echo=TRUE}

#Fluency
Kappa_Fluency <- Kappa(UUT_Fluency)
Kappa_Fluency
confint(Kappa_Fluency)


#Flexibility
Kappa_Flexibility <- Kappa(UUT_Flexibility)
Kappa_Flexibility
confint(Kappa_Flexibility)

```


## Results

The **unweighted** Kappa values demonstrate that the inter-rater reliability is almost perfect (>.90 for both). Importantly, it was expected that it would be slightly higher for fluency as this is more 'straightforward', whereas flexibility is more dependent on the rater. 95% CIs are .95-1 for fluency and .88-.95 for flexibility.



